---
title: "Tidy Data with tidyr"
author: "Vineed"
date: "6/2/2020"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

### Introduction 
```{r}
library(tidyverse)

table1 %>% mutate(
  rate = cases/population*10000
)

#compute total cases per year and rate per year

table1 %>% group_by(year) %>% 
  summarise(total_cases = sum(cases),
            
            total_population = sum(population),
            rate = total_cases/total_population
            )
 
#total cases
table1 %>% count(year, wt=cases)

#visualizing the data
table1 %>% ggplot(aes(x = year, y = cases))+
  geom_line(aes(group = country, color = country))+
  geom_point(aes(color = country))


```
### Exercises

```{r}

#Rate from Table 2
t2_cases <- table2 %>% filter(type == "cases") %>%
  rename(cases = count)
t2_population <- table2 %>% filter(type =="population") %>% 
  rename(population = count)

req_table <- tibble(
  country = t2_cases$country,
  year = t2_cases$year,
  cases = t2_cases$cases,
  population = t2_population$population
) %>% arrange(country,year)
 
req_table %>% mutate(rate = cases/population*1000) 

table4a
table4b

tibble(
  country = table4a$country,
  year = 1999,
  cases = table4a$'1999',
  population = table4b$'1999'
)

#3
table2 %>% filter(type =="cases") %>% 
  ggplot(aes(year, count))+
  geom_line(aes(group = country))+
  geom_point(aes(color = country))
```

## Spreading and Gathering

```{r}
table4a
table4a %>% gather("1999","2000", key = "year", value = "cases") %>% 
  arrange(country, year)

table4b %>% gather("1999","2000",key = "year", value = "population") %>% 
  arrange(country,year)

#to combine, left join needs to be used - more in chapter 10

tidy4a <- table4a %>% gather("1999","2000", key = "year", value = "case")
tidy4b <- table4b %>% gather("1999", "2000", key = "year", value = "population")
df1 <- left_join(tidy4a,tidy4b)
df2 <- right_join(tidy4a,tidy4b) %>% arrange(year)
identical(df1,df2)


```
### Gathering

```{r}
table2
spread(table2, key =type, value = count)

```
### Exercises

#1
```{r}
stocks <- tibble(
  year = c(2015, 2015, 2016, 2016),
  half = c(1,2,1,2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks
stocks %>% spread(year, return) %>% 
  gather("year","return", '2015','2016', convert = T) #if convert not used, then the year will be character. 
```
#2
```{r}
table4a %>% gather("1999", "2000", key = "year", value = "cases")

```
#3
```{r}
people <- tribble(
  ~name,~key,~value,
  "Philip","age",45,
  "Philip","height",186,
  "Philip", "age",50,
  "Jessica", "age", 37,
  "Jessica", "height", 156
)

#people %>% spread(key, value)

people2 <- people %>% group_by(name,key) %>% mutate(obs= row_number())

people2 %>% spread(key,value)

```
#4
```{r}
(preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",NA, 10,
  "no",20,12
))

preg %>% gather(male,female, key = gender, value = infected) %>% 
  select(gender, everything()) 
```

### Seperating and Pull

```{r}
table3 %>% separate(rate, into = c("cases", "population")) #splits non-alphanumeric character

table3 %>% separate(rate, into = c("cases", "population"), sep = "/") #explicit mention of charater for splitting

table3 %>%  separate(rate, into = c("cases", "population"), convert = T) #convert to required type

#passing arguments will split at the location (+ve from far left, -ve from far right)

table5 <- table3 %>% separate(year, into = c("millenium","century", "year"), sep = c(1,2), convert = T)

table5
 ```

#### Unite

```{r}
unite(table5, year, "millenium", century, year) #in quotes or seperate.
#for no arguments between joined variables :-
unite(table5, year,"millenium", century, year, sep = "") #use quotes when the varible names are weird

```
### Exercises

#1
```{r}
tibble(x = c("a,b,c","d,e,f,g","h,i,j")) %>% 
  separate(x, c("one", "two","three"))

tibble(x = c("a,b,c","d,e,f,g","h,i,j")) %>% 
  separate(x, c("one", "two","three"), extra = "drop")

tibble(x = c("a,b,c","d,e,f,g","h,i,j")) %>% 
  separate(x, c("one", "two","three"), extra = "merge")

#fill

tibble(x = c("a,b,c","d,e","h,i,j")) %>% 
  separate(x, c("one", "two","three"))

tibble(x = c("a,b,c","d,e","h,i,j")) %>% 
  separate(x, c("one", "two","three"), fill = "left")

tibble(x = c("a,b,c","d,e","h,i,j")) %>% 
  separate(x, c("one", "two","three"), fill = "right")


```

#2
```{r}

#Separate by separator
tibble(x = c("X_1", "X_2", "AA_1", "AA_2")) %>%
   separate(x, c("variable", "into"), sep = "_")

#Separate by location/position
tibble(x = c("X1", "X2", "Y1", "Y2")) %>%
  separate(x, c("variable", "into"), sep = c(1))

#Separate by group(regex)
tibble(x = c("X_1", "X_2", "AA_1", "AA_2")) %>%
  extract(x, c("variable", "id"), regex = "([A-Z])_([0-9])")

tibble(x = c("X1", "X2", "Y1", "Y2")) %>%
  extract(x, c("variable", "id"), regex = "([A-Z])([0-9])")

tibble(x = c("XXXX1", "X20", "YA1", "YA20")) %>%
  extract(x, c("variable", "id"), regex = "([A-Z]+)([0-9]+)")
```

### Missing Values

```{r}
(stocks <- tibble(
  year = c(2015,2015,2015,2015,2016,2016,2016),
  qtr = c(1,2,3,4,2,3,4),
  return = c(1.88, 0.59, 0.35, NA,0.92, 0.17, 2.66)
))

stocks %>% spread(year, return)

stocks %>% spread(year, return) %>% 
  gather('2015','2016', key = year, value = return)

stocks %>% complete(year, qtr) #Complete takes out all combination of the parameters and bring out implicit missing notes

(treatment <- tribble(
  ~person, ~treatment, ~response,
  "Derrick Whitmore", 1, 7,
  NA,2,10,
  NA,3,9,
  "Katherine Burke", 1, 4
))

treatment %>% fill(person)
treatment %>% fill(person, .direction = "up")



```

## Case Study

```{r}
View(tidyr::who)

who1 <- tidyr::who %>% gather(
  5:60, key = "key", value = "cases", na.rm = T
)

n_distinct(who1$key)

who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key,"newrel", "new_rel"))

View(who2)

who3 <- who2 %>% separate(key, c("new","type","sexage"), sep = "_")

n_distinct(who3$new)

who4 <- who3 %>% select(-new, -iso2, -iso3)

who5 <- who4 %>% separate(sexage, c("sex", "age"), sep = 1)

n_distinct(who5$age)

#replacing with a complex pipe :-

who %>% gather(5:60, key = "code", value = "cases", na.rm = T) %>% 
  mutate(code= stringr::str_replace(code,"newrel","new_rel")) %>% 
  separate(code, c("new","type","sexage"), sep ="_") %>% 
  separate(sexage, c("sex", "age"), sep = 1) %>% 
  select(-iso2,-iso3,-new)


select(who3, country,iso2, iso3) %>% 
  distinct() %>% 
  group_by(country) %>% 
  filter(n()>1)
             
