---
title: "Data Transformation using dplyr, Part 1"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(nycflights13)
library(tidyverse)
```

### Filter Rows with filter()

* Alt + (minus) for <-

```{r}
# Pick observations by their values - filter(data frame, what to do with the data frame, using variables) --> results in another data.frame
jan1 <- filter(flights, month == 1, day == 1) #results in a data frame with Jan 1st as date

(dec25 <- filter(flights, month == 12, day ==25)) #to display results, wrap in paranthesis

#comparison operators are < , <= , > , >= , == , !=

near(sqrt(2)^2, 2)# instead of sqrt(2)^2 == 2

#Boolean operators : & is 'and', | is 'or' and ! is 'not'

nov_dec <- filter(flights, month %in% c(11,12)) #instead of month == 11|12
#nov_dec <-  filter(flights, month ==11 | month == 12) #both evaluate to the same

#Finding flights not delayed more than 2 hours in arrival/departure
filter(flights, !(arr_delay >120 | dep_delay >120))
#applying Demorgan's Rule
filter(flights, !arr_delay >120 & !dep_delay >120)


#filter() by default discards NA values. 

df <- tibble(x = c(1,NA, 3))
filter(df, x >1)
filter(df, is.na(x)|x>1)
```

### Exercises
1
```{r}
#a
filter(flights, arr_delay >=120)
#b
filter(flights, dest %in% c("IAH", "HOU"))
#c
filter(flights, carrier %in% c('UA', 'AA' , 'DL'))
#d
filter(flights, month %in% c(7,8,9))
#e
filter(flights, arr_delay > 120, dep_delay <=0)
#f
filter(flights, dep_delay > 60, arr_delay < dep_delay - 30)
#g
filter(flights, (dep_time >2200 | dep_time <600) )
```

2
```{r}
filter(flights, !between(dep_time,600,2200))
```

3
```{r}
nrow(filter(flights, is.na(dep_time)))
filter(flights, is.na(dep_time))

```

