---
title: "Transformation with dplyr, Part 4"
author: "Vineed"
date: "5/18/2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Add new varibles with mutate()
  + adds new column at the end of the dataset

```{r}

library(tidyverse)
flights_sml <- select(flights,
                      year:day, ends_with("delay"),
                      distance,
                      air_time)

mutate(flights_sml, 
       gain = arr_delay - dep_delay,
       speed = distance / (air_time / 60)
       )

mutate(flights_sml, gain = arr_delay - dep_delay,
       hours = air_time/60,
       gain_per_hour = gain/hours
       )
```

> if we only need the new values, use transmute() instead of mutate()

```{r}
transmute(flights,gain = arr_delay - dep_delay,
       hours = air_time/60,
       gain_per_hour = gain/hours
       )
```

### Useful Creation Functions

  + Arithmetic OPerators
  + Modular Arithmetic (%/% - integer division and %% - remainder)
  + Logs - log(), log2(), log10()
  + Offsets - lead(), lag()
  + cumulative aggregates - cummean(), cumsum(), cumprod(), cummin(), cummax()
  + logical comparisons - >,<, <=, >=, ==, !=
  + ranking - min_rank(), percent_rank(), row_number(), cume_dist(),dense_rank()

```{r}
transmute(flights, hours = dep_time %/% 100,
          minutes = dep_time %% 100)
```

## Exercises
1
```{r}
mutate(flights, 
       dep_minutes = dep_time %/%100 * 60 + dep_time%%100,
       sched_dep_minutes = sched_dep_time %/%100 * 60 + sched_dep_time%%100
       
       )
```
2
```{r}

new_data <- transmute(flights, air_time, 
          arr_minute = (arr_time %/% 100 * 60 + arr_time%%100) %% 1440 , 
          dep_minute = (dep_time %/% 100 * 60 + dep_time%%100) %% 1440, 
          arr_minute - dep_minute)

transmute()

arrange(filter(new_data, arr_minute < dep_minute ), desc(dep_minute))


```
3
```{r}
transmute( dep_time, sched_dep_time, dep_delay)
ggplot(flights)+
  geom_bar(aes(x = dep_time))

```
4.
```{r}
mutate(flights, min_rank(sched_arr_time - arr_time))

```

